Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach

### 2013-09-08 inst/doc/nonparametric-bayes.tex
update chunk options

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\DefineShortVerb[commandchars=\\\{\}]{\|}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

add references to {[}@Fischer2009, @Polasky2011{]} as persective on
combining complex / critical dynamics + control theory

@@

Decision-theoretic (or optimal control) tools require a model that can
assign probabilities of future states (e.g.~stock size of a fishery)
given the current state and a proposed action (e.g.~fishing harvest or
effort). The decision maker then seeks to determining the course of
actions (also referred to as the policy) that maximizes the expected
value of some objective function (such as net present value derived from
the resource over time. (The approach can be adapted to permit
alternatives to maximizing expectation, such as minimizing the maximum
cost or damage that might be incurred. See @Polasky2011). Management
frequently faces a sequential decision-making problem -- after selecting
an action, the decision-maker may receive new information about the
current state and must again choose an appropriate action --
i.e.~setting the harvest limits each year based on stock assessments the
year prior.

More historical context, e.g. {[}@Reed1979, @Mangel1985{]}

@@

{[}@Mangel1985; @Schapaugh2013{]}

@@

{[}@Williams2001; @Athanassoglou2012{]}.

@@

The nature of decision-making problems provides a convenient way to
compare models. Rather than compare models in terms of best fit to data
or fret over the appropriate penalty for model complexity, model
performance is defined in the concrete terms of the decision-maker's
objective function, which we will take as given. (Much argument can be
made over the `correct' objective function, e.g.~how to account for the
social value of fish left in the sea vs.~the commercial value of fish
harvested; see @Halpern2013 for further discussion of this issue.
Alternatively, we can always compare model performance across multiple
potential objective functions.) The decision-maker does not necessarily
need a model that provides the best mechanistic understanding or the
best long-term outcome, but rather the one that best estimates the
probabilities of being in different states as a result of the possible
actions.

@@

In our example, we focus on the problem in which a manager must set the
harvest level for a marine fishery each year to maximize the net present
value of the resource, given an estimated stock size from the year
before. The economic value and ecological concern have made marine
fisheries the crucible for much of the founding work {[}@Gordon1954;
@Reed1979; @May1979; @Ludwig1982{]} in managing ecosystems under
uncertainty. Global trends {[}@Worm2006{]} and controversy
{[}@Hilborn2007; @Worm2009{]} have made understanding these challenges
all the more pressing.

@@

Given parameters for the function $f$ and probability distribution $Z$,
along with a given economic model determining the price/profit
$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
and observed stock $X_t$. This problem can be solved exactly for
discretized values of stock $X$ and policy $h$ using stochastic dynamic
programming (SDP) {[}@Mangel1985{]}. Problems of this sort underpin much
marine fisheries management today.

A crux of this approach is correctly specifying the functional form of
$f$, along with its parameters. The standard approach uses one of a
handful of common parametric models representing the stock-recruitment
relationship, usually after estimating the model parameters from any
available existing data. Uncertainty in the parameter estimates can be
estimated and integrated over to determine the optimal policy under
under uncertainty {[}@Mangel1985; @Schapaugh2013{]}. Uncertainty in the
model structure itself can only be addressed in this approach by
hypothesizing alternative model structures, and then performing some
model choice or model averaging {[}@Williams2001; @Athanassoglou2012{]}.

@@

Concerns over the potential for tipping points in ecological dynamics
{[}@Scheffer2001{]} highlight the dangers of uncertainty in ecological
management and pose a substantial challenge to existing
decision-theoretic approaches {[}@Brozovic2011{]}. Because intervention
is often too late after a tipping point has been crossed (but see
@Hughes2013), management is most often concerned with avoiding
potentially catastrophic tipping points before any data is available at
or following a transition that would more clearly reveal these regime
shift dynamics {[}e.g. @Bestelmeyer2012{]}.

To illustrate the value of the non-parametric Bayesian approach to
management, we focus on example of a system containing such a tipping
point whose dynamics can still be described by a simple, one-dimensional
parametric model.\\We will focus on a simple parametric model for a
single species {[}derived from fist principles by @Allen2005a{]} as our
underlying ``reality''.

@@

As a low-dimensional system completely described by three parameters,
this scenario should if anything be favorable to a parametric-based
approach. This model contains an Allee effect, or tipping point, below
which the population is not self-sustaining and shrinks to zero
{[}@Courchamp2008{]}.

@@

We generate initial observational data under the model described in Eq 1
for $T_{\textrm{obs}}=40$ time steps, under a given arbitrary sequence
of harvest intensities, $h_t$. We consider the case in which most of the
data comes from a limited region of state space (e.g.~near a stable
equilibrium), leaving us without observations of the population dynamics
at very low levels which would be useful in discrimating between
recruitment curves {[}@{]} or demonstrating the existence of a tipping
point {[}@Scheffer2001{]}.\\Using data simulated from a specified model
rather than empirical data permits the comparison against the true
underlying dynamics, setting a bar for the optimal performance possible.

@@

The use of Gaussian process (GP) regression (or ``kriging'' in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems {[}@Kocijan2005{]} and
introduced in the ecological modeling and fisheries management by
@Munch2005. An accessible and thorough introduction to the formulation
and use of GPs can be found in @Rasmussen2006.

@@

The posterior distribution for the hyper-parameters of the Gaussian
process model are estimated by Metropolis-Hastings algorithm, again with
details and code provided in the Appendix. @Rasmussen2006 provides an
excellent general introduction to Gaussian Processes and @Munch2005
first discusses their application in the context of population dynamics
models such as fisheries stock-recruitment relationships.

@@

Once the posterior Gaussian process (GP) has been estimated {[}e.g.~see
@Munch2005{]}, it is necessary to adapt it in place of the parametric
equation for the stochastic dynamic programming (SDP) solution {[}see
@Mangel1988 for a detailed description of parametric SDP methods{]} to
the optimal policy. The essence of the idea is straight forward -- we
will use the estimated GP in place of the parametric growth function to
determine the stochastic transition matrix on which the SDP calculations
are based. The SDP is solved in a discretized state space -- both the
continuously valued population densities $X$ and harvest quotas $h$ are
first mapped to a bounded, discrete grid. (For simplicity we will
consider a uniform grid, though for either parametric or GP-based SDP it
is often advantageous to use a non-uniform discretization such as a
basis function representation, e.g.~see {[}@Deisenroth2009{]}).

@@

@@

The resulting optimal management strategy based on each of the inferred
models is shown in Figure 2, against the optimal strategy given the true
underlying dynamics. Policies are shown in terms of target escapement,
$S_t$. Under models such as this a constant escapement policy is
expected to be optimal {[}@Reed1979{]}, whereby population levels below
a certain size $S$ are unharvested, while above that size the harvest
strategy aims to return the population to $S$, resulting in the
hockey-stick shaped policies shown.

@@

@@

Non-parametric Bayesian methods have received far too little attention
in ecological modeling efforts that are aimed at improved conservation
planning and decision making support. Such approaches may be
particularly useful when the available data is restricted to a limited
area of state-space, which can lead parametric models to underestimate
the uncertainty in dynamics at population levels (states) which have not
been observed. One reason for the relative absence of nonparametric
approaches in the natural resource management context may be the lack of
existing approaches for adapting the non-parametric Bayesian models
previously proposed {[}@Munch2005{]} to a decision-theoretic framework.
Adapting a non-parametric approach requires modification of existing
methods for decision theory. We have illustrated how this might be done
for a classic stochastic dynamic programming problem, opening the door
for substantial further research into how these applications might be
improved.

Though simple mechanistically motivated models may be best for the
latter case {[}@Cuddington2013; @Geritz2012{]}, such models can be not
only inaccurate but misleading in quantitative decision making. This
arises for two reasons:

@@

The complexity of ecological interactions and a lack of data contribute
greatly to both of the problems. This concern is particularly acute in
case of ecological tipping points {[}@Barnosky2012; @Scheffer2001{]},
which arise from feedbacks common in ecological systems and can lead to
a sudden catastrophic transition to an undesirable state. In general we
do not know where such tipping points are unless we have already
observed the transition, in which case it is frequently too late to
respond. Though there may exist early warning signals for certain kinds
of these transitions that are driven by slow changes
{[}@Scheffer2009{]}, we do not know when, where, or how to apply them to
the decision making context more generally {[}@Boettiger2013{]}.

@@

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\DefineShortVerb[commandchars=\\\{\}]{\|}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

add references to (@Fischer2009, Polasky et al. 2011) as persective on
combining complex / critical dynamics + control theory

@@

Decision-theoretic (or optimal control) tools require a model that can
assign probabilities of future states (e.g.~stock size of a fishery)
given the current state and a proposed action (e.g.~fishing harvest or
effort). The decision maker then seeks to determining the course of
actions (also referred to as the policy) that maximizes the expected
value of some objective function (such as net present value derived from
the resource over time. (The approach can be adapted to permit
alternatives to maximizing expectation, such as minimizing the maximum
cost or damage that might be incurred. See Polasky et al. (2011)).
Management frequently faces a sequential decision-making problem --
after selecting an action, the decision-maker may receive new
information about the current state and must again choose an appropriate
action -- i.e.~setting the harvest limits each year based on stock
assessments the year prior.

More historical context, e.g. (Reed 1979, )

@@

(Mangel 1985; Schapaugh and Tyre 2013)

@@

(B. K. Williams 2001; Athanassoglou and Xepapadeas 2012).

@@

The nature of decision-making problems provides a convenient way to
compare models. Rather than compare models in terms of best fit to data
or fret over the appropriate penalty for model complexity, model
performance is defined in the concrete terms of the decision-maker's
objective function, which we will take as given. (Much argument can be
made over the `correct' objective function, e.g.~how to account for the
social value of fish left in the sea vs.~the commercial value of fish
harvested; see Halpern et al. (2013) for further discussion of this
issue. Alternatively, we can always compare model performance across
multiple potential objective functions.) The decision-maker does not
necessarily need a model that provides the best mechanistic
understanding or the best long-term outcome, but rather the one that
best estimates the probabilities of being in different states as a
result of the possible actions.

@@

In our example, we focus on the problem in which a manager must set the
harvest level for a marine fishery each year to maximize the net present
value of the resource, given an estimated stock size from the year
before. The economic value and ecological concern have made marine
fisheries the crucible for much of the founding work (Gordon 1954; Reed
1979; May et al. 1979; Ludwig and Walters 1982) in managing ecosystems
under uncertainty. Global trends (Worm et al. 2006) and controversy
(Hilborn 2007; Worm et al. 2009) have made understanding these
challenges all the more pressing.

@@

Given parameters for the function $f$ and probability distribution $Z$,
along with a given economic model determining the price/profit
$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
and observed stock $X_t$. This problem can be solved exactly for
discretized values of stock $X$ and policy $h$ using stochastic dynamic
programming (SDP) (Mangel 1985). Problems of this sort underpin much
marine fisheries management today.

A crux of this approach is correctly specifying the functional form of
$f$, along with its parameters. The standard approach uses one of a
handful of common parametric models representing the stock-recruitment
relationship, usually after estimating the model parameters from any
available existing data. Uncertainty in the parameter estimates can be
estimated and integrated over to determine the optimal policy under
under uncertainty (Mangel 1985; Schapaugh and Tyre 2013). Uncertainty in
the model structure itself can only be addressed in this approach by
hypothesizing alternative model structures, and then performing some
model choice or model averaging (B. K. Williams 2001; Athanassoglou and
Xepapadeas 2012).

@@

Concerns over the potential for tipping points in ecological dynamics
(Scheffer et al. 2001) highlight the dangers of uncertainty in
ecological management and pose a substantial challenge to existing
decision-theoretic approaches (Brozović and Schlenker 2011). Because
intervention is often too late after a tipping point has been crossed
(but see Hughes et al. (2013)), management is most often concerned with
avoiding potentially catastrophic tipping points before any data is
available at or following a transition that would more clearly reveal
these regime shift dynamics (e.g. Bestelmeyer et al. 2012).

To illustrate the value of the non-parametric Bayesian approach to
management, we focus on example of a system containing such a tipping
point whose dynamics can still be described by a simple, one-dimensional
parametric model.\\We will focus on a simple parametric model for a
single species (derived from fist principles by Allen et al. 2005) as
our underlying ``reality''.

@@

As a low-dimensional system completely described by three parameters,
this scenario should if anything be favorable to a parametric-based
approach. This model contains an Allee effect, or tipping point, below
which the population is not self-sustaining and shrinks to zero
(Courchamp, Berec, and Gascoigne 2008).

@@

We generate initial observational data under the model described in Eq 1
for $T_{\textrm{obs}}=40$ time steps, under a given arbitrary sequence
of harvest intensities, $h_t$. We consider the case in which most of the
data comes from a limited region of state space (e.g.~near a stable
equilibrium), leaving us without observations of the population dynamics
at very low levels which would be useful in discrimating between
recruitment curves {[}@{]} or demonstrating the existence of a tipping
point (Scheffer et al. 2001).\\Using data simulated from a specified
model rather than empirical data permits the comparison against the true
underlying dynamics, setting a bar for the optimal performance possible.

@@

The use of Gaussian process (GP) regression (or ``kriging'' in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems (Kocijan et al. 2005) and
introduced in the ecological modeling and fisheries management by Munch
et al. (2005). An accessible and thorough introduction to the
formulation and use of GPs can be found in Rasmussen and Williams
(2006).

@@

The posterior distribution for the hyper-parameters of the Gaussian
process model are estimated by Metropolis-Hastings algorithm, again with
details and code provided in the Appendix. Rasmussen and Williams (2006)
provides an excellent general introduction to Gaussian Processes and
Munch et al. (2005) first discusses their application in the context of
population dynamics models such as fisheries stock-recruitment
relationships.

@@

Once the posterior Gaussian process (GP) has been estimated (e.g.~see
Munch et al. 2005), it is necessary to adapt it in place of the
parametric equation for the stochastic dynamic programming (SDP)
solution (see Mangel and Clark 1988 for a detailed description of
parametric SDP methods) to the optimal policy. The essence of the idea
is straight forward -- we will use the estimated GP in place of the
parametric growth function to determine the stochastic transition matrix
on which the SDP calculations are based. The SDP is solved in a
discretized state space -- both the continuously valued population
densities $X$ and harvest quotas $h$ are first mapped to a bounded,
discrete grid. (For simplicity we will consider a uniform grid, though
for either parametric or GP-based SDP it is often advantageous to use a
non-uniform discretization such as a basis function representation,
e.g.~see (Deisenroth, Rasmussen, and Peters 2009)).

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{require}\NormalTok{(MASS)}
\NormalTok{step_ahead <- function(x, f, p)\{}
  \NormalTok{h = }\DecValTok{0}
  \NormalTok{x_predict <- }\KeywordTok{sapply}\NormalTok{(x, f, h, p)}
  \NormalTok{n <- }\KeywordTok{length}\NormalTok{(x_predict) - }\DecValTok{1}
  \NormalTok{y <- }\KeywordTok{c}\NormalTok{(x[}\DecValTok{1}\NormalTok{], x_predict[}\DecValTok{1}\NormalTok{:n])}
  \NormalTok{y}
\NormalTok{\}}
\NormalTok{step_ahead_posteriors <- function(x)\{}
\NormalTok{gp_f_at_obs <- }\KeywordTok{gp_predict}\NormalTok{(gp, x, }\DataTypeTok{burnin=}\FloatTok{1e4}\NormalTok{, }\DataTypeTok{thin=}\DecValTok{300}\NormalTok{)}
\NormalTok{df_post <- }\KeywordTok{melt}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(}\KeywordTok{sample}\NormalTok{(}\DecValTok{100}\NormalTok{), }
  \NormalTok{function(i)\{}
    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{time =} \DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(x), }\DataTypeTok{stock =} \NormalTok{x, }
                \DataTypeTok{GP =} \KeywordTok{mvrnorm}\NormalTok{(}\DecValTok{1}\NormalTok{, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),}
                \DataTypeTok{True =} \KeywordTok{step_ahead}\NormalTok{(x,f,p),  }
                \DataTypeTok{MLE =} \KeywordTok{step_ahead}\NormalTok{(x,f,est$p), }
                \DataTypeTok{Allen =} \KeywordTok{step_ahead}\NormalTok{(x, allen_f, pardist[i,]), }
                \DataTypeTok{Ricker =} \KeywordTok{step_ahead}\NormalTok{(x, ricker_f, ricker_pardist[i,]), }
                \DataTypeTok{Myers =} \KeywordTok{step_ahead}\NormalTok{(x, myers_f, myers_pardist[i,]))}
  \NormalTok{\}), }\DataTypeTok{id=}\KeywordTok{c}\NormalTok{(}\StringTok{"time"}\NormalTok{, }\StringTok{"stock"}\NormalTok{))}
\NormalTok{\}}

\NormalTok{df_post <- }\KeywordTok{step_ahead_posteriors}\NormalTok{(x)}

\KeywordTok{ggplot}\NormalTok{(df_post) + }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, stock)) + }
  \KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, value, }\DataTypeTok{col=}\NormalTok{variable, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(L1,variable)), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) + }
  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{))) }
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-Figureb.pdf}
\caption{plot of chunk Figureb}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{policies <- }\KeywordTok{melt}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{stock=}\NormalTok{x_grid, }\KeywordTok{sapply}\NormalTok{(OPT, function(x) x_grid[x])), }\DataTypeTok{id=}\StringTok{"stock"}\NormalTok{)}
\KeywordTok{names}\NormalTok{(policies) <- }\KeywordTok{c}\NormalTok{(}\StringTok{"stock"}\NormalTok{, }\StringTok{"method"}\NormalTok{, }\StringTok{"value"}\NormalTok{)}

\KeywordTok{ggplot}\NormalTok{(policies, }\KeywordTok{aes}\NormalTok{(stock, stock - value, }\DataTypeTok{color=}\NormalTok{method)) +}
  \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{lwd=}\FloatTok{1.2}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) + }\KeywordTok{xlab}\NormalTok{(}\StringTok{"stock size"}\NormalTok{) + }\KeywordTok{ylab}\NormalTok{(}\StringTok{"escapement"}\NormalTok{)  +}
  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-Figure2.pdf}
\caption{The steady-state optimal policy (infinite boundary) calculated
under each model. Policies are shown in terms of target escapement,
$S_t$, as under models such as this a constant escapement policy is
expected to be optimal (Reed 1979).}
\end{figure}

The resulting optimal management strategy based on each of the inferred
models is shown in Figure 2, against the optimal strategy given the true
underlying dynamics. Policies are shown in terms of target escapement,
$S_t$. Under models such as this a constant escapement policy is
expected to be optimal (Reed 1979), whereby population levels below a
certain size $S$ are unharvested, while above that size the harvest
strategy aims to return the population to $S$, resulting in the
hockey-stick shaped policies shown.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{(sims_data) + }
  \KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, fishstock, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(reps,method), }\DataTypeTok{color=}\NormalTok{method), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) +}
  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-Figure3.pdf}
\caption{Gaussian process inference outperforms parametric estimates.
Shown are 100 replicate simulations of the stock dynamics (eq 1) under
the policies derived from each of the estimated models, as well as the
policy based on the exact underlying model.}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fig4v1 <- }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) + }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable)) + }
  \KeywordTok{facet_wrap}\NormalTok{(~variable, }\DataTypeTok{scales =} \StringTok{"free_y"}\NormalTok{)  + }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{) +}
  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{) + }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey) }\CommentTok{# density plots fail when delta fn}

\NormalTok{fig4v2 <- }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) + }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable), }\DataTypeTok{binwidth=}\FloatTok{0.1}\NormalTok{) + }
  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}

\NormalTok{fig4v3 <- }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value, }\DataTypeTok{fill=}\NormalTok{variable, }\DataTypeTok{color=}\NormalTok{variable)) + }\CommentTok{# density plots fail when delta fn}
  \KeywordTok{stat_density}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{y=}\NormalTok{..density..), }\DataTypeTok{position=}\StringTok{"stack"}\NormalTok{, }\DataTypeTok{adjust=}\DecValTok{3}\NormalTok{, }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{9}\NormalTok{) + }
  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)+ }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
\NormalTok{fig4v2}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-Figure41.pdf}
\caption{Histograms of the realized net present value of the fishery
over a range of simulated data and resulting parameter estimates. For
each data set, the three models are estimated as described above. Values
plotted are the averages of a given policy over 100 replicate
simulations. Details and code provided in the supplement.}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fig4v3}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-Figure42.pdf}
\caption{Histograms of the realized net present value of the fishery
over a range of simulated data and resulting parameter estimates. For
each data set, the three models are estimated as described above. Values
plotted are the averages of a given policy over 100 replicate
simulations. Details and code provided in the supplement.}
\end{figure}

@@

Non-parametric Bayesian methods have received far too little attention
in ecological modeling efforts that are aimed at improved conservation
planning and decision making support. Such approaches may be
particularly useful when the available data is restricted to a limited
area of state-space, which can lead parametric models to underestimate
the uncertainty in dynamics at population levels (states) which have not
been observed. One reason for the relative absence of nonparametric
approaches in the natural resource management context may be the lack of
existing approaches for adapting the non-parametric Bayesian models
previously proposed (Munch et al. 2005) to a decision-theoretic
framework. Adapting a non-parametric approach requires modification of
existing methods for decision theory. We have illustrated how this might
be done for a classic stochastic dynamic programming problem, opening
the door for substantial further research into how these applications
might be improved.

Though simple mechanistically motivated models may be best for the
latter case (Cuddington, Fortin, and Gerber 2013; Geritz and Kisdi
2012), such models can be not only inaccurate but misleading in
quantitative decision making. This arises for two reasons:

@@

The complexity of ecological interactions and a lack of data contribute
greatly to both of the problems. This concern is particularly acute in
case of ecological tipping points (Barnosky et al. 2012; Scheffer et al.
2001), which arise from feedbacks common in ecological systems and can
lead to a sudden catastrophic transition to an undesirable state. In
general we do not know where such tipping points are unless we have
already observed the transition, in which case it is frequently too late
to respond. Though there may exist early warning signals for certain
kinds of these transitions that are driven by slow changes (Scheffer et
al. 2009), we do not know when, where, or how to apply them to the
decision making context more generally (Boettiger and Hastings 2013).

@@

Allen, Linda J. S., Jesse F. Fagan, Göran Högnäs, and Henrik Fagerholm.
2005. ``Population extinction in discrete-time stochastic population
models with an Allee effect.'' \emph{Journal of Difference Equations and
Applications} 11 (4-5) (apr): 273--293.
doi:10.1080/10236190412331335373.
\url{http://www.tandfonline.com/doi/abs/10.1080/10236190412331335373}.

Athanassoglou, Stergios, and Anastasios Xepapadeas. 2012. ``Pollution
control with uncertain stock dynamics: When, and how, to be
precautious.'' \emph{Journal of Environmental Economics and Management}
63 (3) (may): 304--320. doi:10.1016/j.jeem.2011.11.001.
\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069611001409}.

Barnosky, Anthony D., Elizabeth a Hadly, Jordi Bascompte, Eric L.
Berlow, James H. Brown, Mikael Fortelius, Wayne M. Getz, et al. 2012.
``Approaching a state shift in Earth's biosphere.'' \emph{Nature} 486
(7401) (jun): 52--58. doi:10.1038/nature11018.
\url{http://www.nature.com/doifinder/10.1038/nature11018}.
